---
title: "Covid-19 Vaccination Rate"
author: "Jinsung Park"
format: gfm
---

# Covid-19 Vaccination Rates

Download the Statewide COVID-19 Vaccines csv file.

Limitations of the data:

Data don't include doses administered by the following federal agencies who received vaccine allocated directly from CDC: Indian Health Service, Veterans Health Administration, Department of Defense, and the Federal Bureau of Prisons.
Zip code areas that include military bases will likely show artificially low vaccination rates



## Reading the file

```{r}
vax <- read.csv("Statewide COVID-19 Vaccines.csv")
head(vax)
```

> Q1. What column details the total number of people fully vaccinated? 

Column 'tot_population' shows the total number of people fully vaccinated in each area.

> Q2. What column details the Zip code tabulation area? 

Column 'zip_code_tabulation_area' contains zip code of each area.

> Q3. What is the earliest date in this dataset? 

```{r}
min(vax$as_of_date)
```

The earliest date in the dataset is Jan 05, 2021.

> Q4. What is the latest date in this dataset? 

```{r}
max(vax$as_of_date)
```

The latest data in the dataset is Nov 15, 2022.


Let's use 'skim()' function from the skimr package to get a quick overview

```{r}
skimr::skim(vax)
```


> Q5. How many numeric columns are in this dataset? 

There are 13 numeric columns in the dataset.

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

'persons_fully_vaccinated' column contains 15440 NA values

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

89% of 'persons_fully_vaccinated' are missing

> Q8. [Optional]: Why might this data be missing?

This data might be missing due to lack of vaccination record performed by federal agency.


## Working with Dates

'as_of_date' column contains dates in the Year-Month-Day format.

'lubridate' package can make life allot easier when dealing with date format.

```{r}
library(lubridate)
```
```{r}
today()
```

This is cool!

By using 'lubridate' we can turn 'as_of_date' column into the usable form.
We need to provide argument first to allow us to do math with date.

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset? 

```{r}
today()-max(vax$as_of_date)
```

It has been 7 days since the last update of the dataset.

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
nrow(table(vax$as_of_date))
```

There are 98 different dates in the dataset.


## Working with Zipcode

We can use the 'zipcodeR' package to make working with these codes easier. 

```{r}

```


## Focus on San Diego Area

We can focus by restricting ourselves first to vax$county == "San Diego"

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

Using dplyr also allow us to subset across multiple criteria

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
nrow(table(sd$zip_code_tabulation_area))
```

There are 107 distinct zip codes listed for San Diego County.


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 

```{r}
which.max(sd$age12_plus_population)
sd[90,2]
```

Zip code with the largest 12+ population value is 92154. 


> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”? 

```{r}
library(dplyr)
sd1115 <- filter(vax, county == "San Diego" & as_of_date == "2022-11-15")
nrow(sd1115)
```
```{r}
sd1115_mean <- mean(sd1115$percent_of_population_fully_vaccinated, na.rm = TRUE)
sd1115_mean
```

Average of fully vaccinated population percentage is 0.738


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?

```{r}
library(ggplot2)
df <- sd1115
p <- ggplot(df, aes(x = percent_of_population_fully_vaccinated)) + geom_histogram(color = "black", fill="white")
p
```


## Focus on UCSD/La Jolla

UC San Diego resides in the 92037 ZIP code area and is listed with an age 5+ population size of 36,144.

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
p <- ggplot(ucsd) +
  aes(x = as_of_date,y = percent_of_population_fully_vaccinated) + geom_point() + geom_line(group=1) + ylim(c(0,1)) +
  labs(x = "Date", y = "Percent Vaccinated")
p
```


## Comparing to similar sized areas

Let’s return to the full dataset and look across every zip code area with a population at least as large as that of 92037 on as_of_date “2022-02-22”.


```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-11-15")

head(vax.36)
```


> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?



```{r}
vax_mean <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
vax_mean
```
```{r}
p2 <- ggplot(ucsd) +
  aes(x = as_of_date,y = percent_of_population_fully_vaccinated) + geom_point() + geom_line(group=1) + ylim(c(0,1)) + ggtitle("Vaccination Rate for La Jolla CA 92109") + geom_hline(yintercept = vax_mean, linetype="dashed", color = "red") +
  labs(x = "Date", y = "Percent Vaccinated")
p2
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```


> Q18. Using ggplot generate a histogram of this data.

```{r}
df <- vax.36
p3 <- ggplot(df, aes(x = percent_of_population_fully_vaccinated)) + geom_histogram(color = "black", fill="white")
p3
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
Both area 92040 and 92109 are below the average for fully vaccinated population percentages for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”.


> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax,age5_plus_population > 36144)

ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x = "Date", y = "Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only Areas with a Popuplation Above 36k are Shown") +
  geom_hline(yintercept = vax_mean, linetype="dashed", color = "red")
```


> Q21. How do you feel about traveling for Thanksgiving Break and meeting for in-person class afterwards?

Pretty safe, as about more than half of the areas seem to be vaccinated above the average.


